A Data-Driven Framework for Neural Field Modelling

### 2010-11-19 ltx/first_order_framework/LaTeX_Manuscript.tex
deleted sigmoid figure
corrected cross referencing

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Fig.~\ref{fig:Figure9} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is again mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Fig.~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Fig.~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field, shown in Fig.~\ref{fig:Figure10}(A), but is missing the higher-frequency detail, which can be see in the error shown in Fig.~\ref{fig:Figure10}(C).

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Fig.~\ref{fig:Figure8}. For Fig.~\ref{fig:Figure8}(A-C) the Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Fig.~\ref{fig:Figure8}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accord with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the widths of the Gaussians forming the kernel are precisely known, the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

Fig.~\ref{fig:Figure12} demonstrates the effect of the sigmoid parameters on the estimation accuracy. The RMSE of the reconstructed connectivity kernel was calculated for various parameters of the sigmoid by varying the slope, $\varsigma$, and the threshold, $v_0$. The other model parameters remained fixed and were set according to the Table~\ref{tab:Parameters}.

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Fig.~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid. For example, when the sigmoid slope is very gentle, most of the neural activity falls in the central, nearly linear region of the sigmoid, thus having minor effects on the estimation procedure. On the other hand, sharp slopes result in most of the activity falling in the saturation region of the sigmoid with adverse effects on the estimation. Similar arguments can be made for variations to the sigmoid threshold.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Fig.~\ref{fig:Figure13} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} states, the observation kernel acts as a spatial filter.

When the observation kernel is too narrow, high frequency components in the neural field are present but under sampled. These high frequency components result in aliasing errors in the estimated field, thus affecting the kernel estimation accuracy. On the other hand, for the very wide observation kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the important frequency components by the sensors, resulting in poor accuracy in the kernel estimate. As indicated in Fig.~\ref{fig:Figure13}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

\subsection{Implications for Experimental Design}
As seen in Fig.~\ref{fig:Figure7} and Fig.~\ref{fig:Figure10}, the results show good estimation accuracy for the connectivity kernels, where the true kernels lie within the 95~\% confidence intervals of the kernel estimates. The estimates of the connectivity kernels show relatively broad distributions over the different realisations, which can be expected considering the system's input is modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, the use of a controlled input stimulus should be considered when applying the estimation algorithm to electrophysiological recordings, where given knowledge of the input and the appropriate model, it is expected that the variance of the parameter estimates would decrease. Even when using a known input, our results imply that we should only expect to learn distributions of parameters, rather than uncover a `true' parameter set. By designing an input signal carefully and using an evoked-potential paradigm where stimuli are repeated many times, recordings could be averaged over a sufficient number of trials (realisations), such that a distribution of the parameter estimates and resulting connectivity kernels can be formed.

@@

{\bf Figure 13. RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig1.pdf} 
\end{center}
\caption{}
\label{fig:Figure1}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig2.pdf} 
\end{center}
\caption{}
\label{fig:Figure2}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig3.pdf}
\end{center}
\caption{} 
\label{fig:Figure3}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig4.pdf}
\end{center}
\caption{}
\label{fig:Figure4}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht] 
\begin{center}
\includegraphics{./Graph/pdf/fig5.pdf}	
\end{center}
\caption{}
\label{fig:Figure5}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig6.pdf}
\end{center}
\caption{}
\label{fig:Figure6}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig7.pdf}
\end{center}
\caption{} 
\label{fig:Figure7}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig8.pdf} 
\end{center}
\caption{} 
\label{fig:Figure9}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig9.pdf}
\end{center}
\caption{}
\label{fig:Figure10}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig10.pdf}
\end{center}
\caption{}
\label{fig:Figure8}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig11.pdf}
\end{center}
\caption{}
\label{fig:Figure12}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig12.pdf}
\end{center}
\caption{}
\label{fig:sigmoidfunctions}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig13.pdf}
\end{center}
\caption{}
\label{fig:Figure13}
\end{figure*}
\clearpage
\newpage
\appendix
\section{Time Discretisation of Neural Field Model}
\label{App:Time Discretization} To form the integro-difference equation neural field model, a time discretisation must be performed. To do this, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

>>>>

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for 150 realisations is shown in Fig.~\ref{fig:Figure10} with the $95~\%$ confidence interval. At each time point, the error is consistently around $0.5$~mV, which is approximately $10.73~\%$ of the field voltage range ([-2.20, 2.46]~mV). The error in the reconstructed field is again mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Fig.~\ref{fig:Figure9}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field, and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Fig.~\ref{fig:Figure9}(B) that the estimated field contains the basic structure of the true field, shown in Fig.~\ref{fig:Figure9}(A), but is missing the higher-frequency detail, which can be see in the error shown in Fig.~\ref{fig:Figure9}(C).

@@

To study the robustness of the estimation algorithm, the performance was evaluated for various connectivity kernel parameters. The reconstructed connectivity kernels from the estimated parameters are shown in Fig.~\ref{fig:Figure10}. For Fig.~\ref{fig:Figure10}(A-C) the Mexican hat connectivity kernels were used. For each subset of parameters, the mean of the reconstructed kernel over 50 realisations is shown as a dotted line and the 95~\% confidence interval is indicated by the grey shaded region. In each case, the actual kernel was within the 95~\% confidence interval of the kernel estimate. The statistics for the kernel parameter estimates for each of the parameter sets are shown in Table~\ref{tab:MexicanHatKernelResults}.

For each of the Mexican hat connectivities, the support of the kernel was assumed to be known. However, in Fig.~\ref{fig:Figure10}(D-F) the kernel support in the estimator differed from the actual data. In these cases the data sets were generated using a single positive (excitatory) Gaussian kernel, as seen by the solid line in each of the subplots of the figure. The estimation algorithm was constructed allowing for the Mexican hat connectivity of the previous examples. Each plot shows that the actual kernel lies within the confidence intervals of the reconstructed estimates for the majority of the kernel space, until the amplitude is relatively low. Table~\ref{tab:GaussianKernelResults} shows the statistics of the parameter estimates. Although the net reconstructed kernel estimates are in good accord with the actual kernels, the individual kernel basis function weights are not in agreement with the actual parameters. This result demonstrates that unless the widths of the Gaussians forming the kernel are precisely known, the interpretation of the results should be based on the net shape of the estimated connectivity kernel and not on the estimated parameter values individually.

@@

Fig.~\ref{fig:Figure11} demonstrates the effect of the sigmoid parameters on the estimation accuracy. The RMSE of the reconstructed connectivity kernel was calculated for various parameters of the sigmoid by varying the slope, $\varsigma$, and the threshold, $v_0$. The other model parameters remained fixed and were set according to the Table~\ref{tab:Parameters}.

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Fig.~\ref{fig:Figure11} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. This is expected since by altering the parameters we are changing the nonlinearity and the saturation region of the sigmoid. For example, when the sigmoid slope is very gentle, most of the neural activity falls in the central, nearly linear region of the sigmoid, thus having minor effects on the estimation procedure. On the other hand, sharp slopes result in most of the activity falling in the saturation region of the sigmoid with adverse effects on the estimation. Similar arguments can be made for variations to the sigmoid threshold.

\subsubsection{Experiment IV: Results for Varying the Sensor Kernel Width}
The accuracy of the estimation algorithm was tested for different observation kernels by varying the width parameter, $\sigma_m$. Fig.~\ref{fig:Figure12} shows the RMSE of the connectivity kernel estimates for the various observation kernel widths. As the theory presented earlier in Section~\ref{SpectralAnalysisSection} states, the observation kernel acts as a spatial filter.

When the observation kernel is too narrow, high frequency components in the neural field are present but under sampled. These high frequency components result in aliasing errors in the estimated field, thus affecting the kernel estimation accuracy. On the other hand, for the very wide observation kernels considered, important neural dynamics are not pick up by the estimated field due to the filtering out of the important frequency components by the sensors, resulting in poor accuracy in the kernel estimate. As indicated in Fig.~\ref{fig:Figure12}, the best results are obtained when the sensor width is chosen to minimize both sources of errors.

@@

\subsection{Implications for Experimental Design}
As seen in Fig.~\ref{fig:Figure7} and Fig.~\ref{fig:Figure9}, the results show good estimation accuracy for the connectivity kernels, where the true kernels lie within the 95~\% confidence intervals of the kernel estimates. The estimates of the connectivity kernels show relatively broad distributions over the different realisations, which can be expected considering the system's input is modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, the use of a controlled input stimulus should be considered when applying the estimation algorithm to electrophysiological recordings, where given knowledge of the input and the appropriate model, it is expected that the variance of the parameter estimates would decrease. Even when using a known input, our results imply that we should only expect to learn distributions of parameters, rather than uncover a `true' parameter set. By designing an input signal carefully and using an evoked-potential paradigm where stimuli are repeated many times, recordings could be averaged over a sufficient number of trials (realisations), such that a distribution of the parameter estimates and resulting connectivity kernels can be formed.

@@

{\bf Figure 13. RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig1.pdf} 
\end{center}
\caption{}
\label{fig:Figure1}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig2.pdf} 
\end{center}
\caption{}
\label{fig:Figure2}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig3.pdf}
\end{center}
\caption{} 
\label{fig:Figure3}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig4.pdf}
\end{center}
\caption{}
\label{fig:Figure4}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht] 
\begin{center}
\includegraphics{./Graph/pdf/fig5.pdf}	
\end{center}
\caption{}
\label{fig:Figure5}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig6.pdf}
\end{center}
\caption{}
\label{fig:Figure6}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig7.pdf}
\end{center}
\caption{} 
\label{fig:Figure7}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig8.pdf} 
\end{center}
\caption{} 
\label{fig:Figure8}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig9.pdf}
\end{center}
\caption{}
\label{fig:Figure9}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig10.pdf}
\end{center}
\caption{}
\label{fig:Figure10}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig11.pdf}
\end{center}
\caption{}
\label{fig:Figure11}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig12.pdf}
\end{center}
\caption{}
\label{fig:Figure12}
\end{figure*}
\clearpage
\newpage
\appendix
\section{Time Discretisation of Neural Field Model}
\label{App:Time Discretization} To form the integro-difference equation neural field model, a time discretisation must be performed. To do this, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.
